\noindent
Thanks to:
\begin{itemize}
    \item My parents, Bill and Katrina McGovern, and my sister Ashleigh for their love, patience, and support throughout my life.
    \item Chris and Kimberly Paterson Hunt, who have been my friends since the beginning of this academic journey.
    \item Cameron and Aline Christensen, for their friendship, beer knowledge, and memes.
    \item George and Carole Leone, for their unending support and guidance throughout my academic career.
    \item Andrew Guenther, for uploading this template.
    \item Roommates Kimberly Paterson Hunt, Eriq Augustine, Chris Hunt, Maria Pikusova, Nicole Martin, and Cody Hunt for the strength of our roommunity.
    \item Alex Dekhtyar for his insight and advisement in the fields of data science, Eurasian indie bands, and novel science fiction and fantasy used book stores.
    \item Chris Kitts, Michael Black, and Jennifer VanderKelen for editing my papers and teaching me everything I needed to know about poop to succeed.
    \item Ilene, Tom, and Brian Jones; Andrew Wang and Daniel Gerrity; and Andrew, Katia, and Finn Guenther for housing me while I presented parts of this thesis at bioinformatics conferences around the country.
    \item The Outrun, Retrowave, and Synthwave genres of music and artists Lazerhawk, FM-84, and Carpenter Brut, whose driving synths and 80's \mbox{\begin{CJK}{UTF8}{min}ａｅｓｔｈｅｔｉｃ\end{CJK}} impelled me to finish this thesis.
    \item And Paul Biggins, for his love and unwavering dedication. 
\end{itemize}
